Precision calculations for cosmic shear

### 2017-02-08 limber2.tex
Finished the discussion of Fig 1 @MK - what is the difference in compute time for full vs 2nd order?

The tomographic shear power spectrum, in analogy to (\ref{eq:C_ell_phi_Pm}), is defined by
%
\begin{equation}
  \left\langle _2\gamma^{}_{\ell m, i} \; {}_2\gamma^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\gamma_{ij}(\ell).
  \label{eq:C_ell_gamma}
\end{equation}
%
For a flat Universe, this is given by
%
\begin{align}
  C^\gamma_{ij}(\ell) = \frac 1 4 \ellbar^2(\ell, 2) \, C^\psi_{ij}(\ell)
                 = & \frac 2 \pi \, \ellbar^2(\ell, 2) \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:C_ell_full}
\end{align}
%
The spherical prefactor for the full projection power spectrum is $p(\ell) =
\ellbar^2(\ell, 2)$, which will be modified under flat-sky and Limber
approximations below.

@@

Considering first the flat-sky cases, the standard first-order Limber approximation, (L1Fl), that was adopted for all pre-2014 CFHTLenS analyses, we find to be 
accurate to better than 10\% for $\ell>3$, converging slowly to the true projection with percent level predicison at $\ell>100$. 
For the extended Limber approximations `hybrid' cases (ExtL1FlHyb and ExtL2FlHyb), despite decreased accuracy
for $\ell < 8$ in comparison to the standard first-order Limber case, the errors with respect to
the true power spectrum decrease much faster, as $\ell^{-2}$, such that percent-level precision is reached at $\ell>15$.  The first-order extended Limber approximation `hybrid' cases (ExtL1FlHyb) was adopted by both \citet{joudaki/etal:2016} and \cite{KiDS-450}\footnote{We confirm that there is a typographical error in equation 4 of \cite{KiDS-450} which should include the extra term of `+0.5' in $\nu(\ell)$ that was incorporated in the cosmological analysis.}

The outlier in the flat-sky cases is the extended Limber approximation (ExtL1Fl) which performs relatively poorly, and
reach 10\% precision only at $\ell > 100$.   To our knowledge this form of the flat-sky approximation has not been used in cosmic shear studies to date.

Including the
spherical prefactor decreases the difference by a factor of a few for $\ell <
5$. Going to second-order yields percent accuracy down to $\ell = 4$.

\mk{Not sure whether the following comments should be left in the draft.} One
notes that the difference between the full projection and second-order Limber
does not scale as expected with $\ell^{-4}$, but rather with $\ell^{-2}$. This
might be due to numerical noise that creates residual terms at the ${\cal
O}(\ell^{-2})$ which do not cancel. Also note that the numerical noise for
$\ell > 20$ is due to numerical integration errors for the full projection.

The flat extended Limber approximations (ExtL1Fl and ExtL2Fl) perform relatively poorly, and
reach 10\% precision only at $\ell > 100$.

>>>>

The tomographic shear power spectrum, in analogy to (\ref{eq:C_ell_phi_Pm}), is defined by
%
\begin{equation}
  \left\langle _2\gamma^{}_{\ell m, i} \; {}_2\gamma^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\gamma_{ij}(\ell).
  \label{eq:C_ell_gamma}
\end{equation}
%
For a zero-curvature Universe, this is given by
%
\begin{align}
  C^\gamma_{ij}(\ell) = \frac 1 4 \ellbar^2(\ell, 2) \, C^\psi_{ij}(\ell)
                 = & \frac 2 \pi \, \ellbar^2(\ell, 2) \, \pref^2
                 \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^{\chi_{\rm lim}} \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:C_ell_full}
\end{align}
%
The spherical prefactor for the full projection power spectrum is $p(\ell) =
\ellbar^2(\ell, 2)$, which will be modified under flat-sky and Limber
approximations below.

@@

Considering first the flat-sky cases, the standard first-order Limber approximation, (L1Fl), that was adopted for all pre-2014 CFHTLenS analyses, we find to be 
accurate to better than 10\% for $\ell>3$, converging slowly to the true projection with percent level precison at $\ell>100$. 
For the extended Limber approximations `hybrid' cases (ExtL1FlHyb and ExtL2FlHyb), despite decreased accuracy
for $\ell < 8$ in comparison to the standard first-order Limber case, the errors with respect to
the true power spectrum decrease much faster, as $\ell^{-2}$, such that percent-level precision is reached at $\ell>15$.  The first-order extended Limber approximation `hybrid' cases (ExtL1FlHyb) was adopted by both \citet{joudaki/etal:2016} and \cite{KiDS-450}\footnote{We confirm that there is a typographical error in equation 4 of \cite{KiDS-450} which should include the extra term of `+0.5' in $\nu(\ell)$ that was incorporated in the cosmological analysis.}.

The outlier in the flat-sky cases is the extended Limber approximation (ExtL1Fl) which performs relatively poorly, and
reaches 10\% precision only at $\ell > 100$.   To our knowledge this form of the flat-sky approximation has not been used in any cosmic shear studies to date, and should not be used in any future studies given these results.  The poor behaviour of this case, in comparison with the `hybrid' case, (e.g ExtL1FlHyb) can be understood by considering Taylor expansions of the different pre-factors.  The spherical pre-factor,
\begin{equation}
p(\ell) = \frac{(\ell+2)(\ell+1)\ell(\ell-1)}{(\ell + 0.5)^4} = 1-{5\over 2\ell^2} +{\cal O}(\ell^{-3}) \, .
\end{equation}
can be compared to the flat-sky extended Limber pre-factor
\begin{equation}
p(\ell) = \ell^4 / (\ell + 0.5)^4 = 1-{2\over \ell}+{5\over 2\ell^2} +{\cal O}(\ell^{-3}) \, , 
\end{equation} 
showing it to be more deviant from the full spherical solution, than the `hybrid' $p(\ell) = 1$ case.

Considering now the spherical-sky cases, we find that including the spherical pre-factor decreases the difference between the extended Limber approximated cases (ExtL1Sph and ExtL2Sph) and the full spherical solution by a factor of a few for $\ell < 5$.   We find that using the spherical-sky second-order extended Limber approximation yields percent accuracy down to $\ell > 3$.  The numerical calculation of the second-order extended Limber approximation is a factor of \ch{X} faster than the calculation of the full spherical solution.   We note that the noise seen in the right panel of Fig.~\ref{fig:Cl_cases} is due to numerical noise arising from numerical integration errors the full spherical solution calculation for
$\ell > 20$.  Increasing numerical precision would increase the calculation time still further, relative to the  second-order extended Limber calculation.

%\mk{Not sure whether the following comments should be left in the draft.} One
%notes that the difference between the full projection and second-order Limber
%does not scale as expected with $\ell^{-4}$, but rather with $\ell^{-2}$. %
%This
%might be due to numerical noise that creates residual terms at the ${\cal
%O}(\ell^{-2})$ which do not cancel. Also note that the numerical noise for
%$\ell > 20$ is due to numerical integration errors for the full projection.
